---
title: "Exam 2022-06-03"
author: "Mohamed Ali - Mohal954"
date: "2023-05-22"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,results="markup")
library(ggplot2)
```

First we load all the data by running the R file givin in the exam
```{r, echo=T}
rm(list = ls())
source("ExamData.R")
set.seed(1)
```

# Question 1  

In this question we have $\alpha$ = 2326 and $\beta$ = 7 We use the rgamma function to generate the 10000 draws by specificied n ) 10000 and And shape = $\alpha$ and Scale = $\beta$.


```{r q1b, echo=T}
theta<-data.frame(x=rgamma(10000,2326,7))
```

### Plot of Theta distribution

You can also embed plots, for example:

```{r q1b1, echo=T}
ggplot() +
  geom_histogram(data=theta,aes(x = x,y=..density..),linetype=1,
                 fill='#14213D')+
  geom_density(data=theta,aes(x=x), color='#FCA311', size=1,
               fill='#FCA311',alpha=.5)+
  labs(title = "Posterior predictive distribution",
       subtitle = " of the demanded quantity in the next month",
       x = "Demanded",
       y = "Density") + theme_classic()
```

Now we find the draws of q6 by plugin the values of theta in rpios function since ur data model follows the Poisson dist. then we find the mean and compare it with the expected value we have 350.

```{r q1b2, echo=T}
mean(rpois(10000,theta$x)>350)
```

### Utility Function 

Given the utility function we want to simulate values to find the optimal a which in returen tell us how much of products we should keep in the stock, in another word the value of a that masimize the utility function.
We first find a sequance of a to plug in the utility function givan the value of Q6.

```{r q1c, echo=T}
a<- seq(1,10000,by=1)
q6<- rpois(10000,theta$x)
res<-data.frame(x=1:10000,y=utility_func(a,q6))
max(res$x,res$y)
ggplot(data=res) +
  geom_line(aes(x = x,y=y),linetype=2)+
  geom_vline(xintercept = which.max(res$y))+
  labs(title = "Posterior predictive distribution",
       subtitle = " of the demanded quantity in the next month",
       x = "Demanded",
       y = "Density") + theme_classic()
```

Which give us the value:

```{r q1c2, echo=T}
which.max(res$y)
```

# Question 2

To answer this question, function BayesLinReg will be used from the exam file, giving the input values for $y, X, \mu_0, \Omega_0, v_0, \sigma2_0$ and nIter. 
From the question we have:

```{r q2a, echo=T}
y<- y
X <- X
mu_0 <- matrix(0,6,1)
Omega_0 <- 10^2*(diag(1,6,6))
v_0 <- 1
sigma2_0 <- 100^2
nIter <- 10000
reg_res<-BayesLinReg(y, X, mu_0, Omega_0, v_0, sigma2_0, nIter)
beta_res<-reg_res$betaSample
sigma2_res<-reg_res$sigma2Sample
```

Now we comupte the 99% CI and mean for all paramters.

```{r q2a1, echo=T}
mean=as.matrix(apply(beta_res, 2, mean))

for (i in 1:length(mean)){
  print(paste0("The mean value of beta_",i-1," equal to "
               ,round(mean[i,1],3)))
}

for (i in 1:length(mean)){
  print(paste0("99% Ci for beta_ ",i))
  print(quantile(beta_res[,i],c(0.01,0.99)))
}
```

The 99% CI for beta_1 shows a lower pound of 6.32 and upper pound of 12.6 which tell us a good induction of the significance of the parameter.

Find the mean and median for sigma

```{r q2b, echo=T}
print(paste0("Mean of Sigma ", mean(sqrt(sigma2_res))))
print(paste0("Median of Sigma ", median(sqrt(sigma2_res))))
```

Now we want to plot the 95% CI for the expected length mu on the grid of values of x_1 for molluscs.

```{r q2c, echo=T}
# First we build our data matrix
x_1= seq(min(X[,2]),max(X[,2]),0.1)
x_2= 27
# Generating the new data
Xnew<-as.matrix(data.frame(intersept=rep(1,length(x_1)),x1=x_1,x12=x_1^2,
           x2=rep(27,length(x_1)),x22=rep(27^2,length(x_1)),
           x1x2=x_1*27))

# finding the mu by multibly the betas by the new x values
mu<-beta_res%*%t(Xnew)

# Finding the CI 
ci<- matrix(0,length(x_1),2)
for (i in 1:length(x_1)){
ci[i,]<-quantile(mu[,i], probs = c(0.05,0.95))
}

# Plotting every thing togther
plot_df<- data.frame(x=x_1)
plot_df<- cbind(plot_df,ci)
colnames(plot_df)<-c("x","lower","upper")

plt <- ggplot(plot_df,aes(x=x)) +
  geom_ribbon(aes(ymin = lower, ymax = upper)
              , alpha = 0.8,fill = "#EDC948")+
  labs(x = 'x_1',title ='. Compute 95% equal tail posterior probability intervals'
       ,color = "Line Legend") +
  scale_color_manual(values = c("#14213D","#59A14F")
                     , labels = c("2","3"))+
  theme(legend.position="bottom")
plt

```
From the above chart we 

Find the simulation for the new Mollusc
```{r q2e, echo=T}
# first we calculate the mu by pluging the values of new Xs and then use the values of betas 
# caluslated in the previous step.
x1<-50
x2<-25
x_new<- c(1,x1,x1^2,x2,x2^2,x1*x1)
mu_new<-beta_res%*%x_new
e<-rnorm(1,0,1/sigma2_0)
y_dist<- data.frame(mu=mu_new,e=e,y=mu_new+e)

ggplot() +
  geom_histogram(data=y_dist,aes(x = y,y=..density..),linetype=1,
                 fill='#14213D')+
  geom_density(data=y_dist,aes(x=y), color='#FCA311', size=1,
               fill='#FCA311',alpha=.5)+
  labs(title = "Posterior predictive distribution of y ",
       subtitle = " y simulation the posterior predictive 
       distribution of y for a new mollusc",
       x = "New Mollusc",
       y = "Density") + theme_classic()
```
Finally find the pr(t(y_rep)>=T(y))
```{r q2f, echo=T}
max_y<-c()
for (i in 1:nIter) {
y_rep<-rnorm(length(y),mu_new[i,],sqrt(sigma2_res[i]))
max_y[i]<-max(y_rep)}
mean(max_y>=max(y))
```
# Question 3

```{r q2e, echo=T}
# To write this function we can get use of the rgamma function in r where we
#just need to specify the alpha and the betas values.
x<-c(0.8,1.1,0.8,0.9,1)
alpha<-3+length(x)
beta<- 4+sum(x^3)
theta<-data.frame(x=rgamma(10000,shape=alpha,scale = 1/beta))
ggplot() +
  geom_histogram(data=theta,aes(x = x,y=..density..),linetype=1,
                 fill='#14213D')+
  geom_density(data=theta,aes(x=x), color='#FCA311', size=1,
               fill='#FCA311',alpha=.5)+
  labs(title = "Posterior predictive distribution of y ",
       subtitle = " y simulation the posterior predictive 
       distribution of y for a new mollusc",
       x = "New Mollusc",
       y = "Density") + theme_classic()
```

