---
title: "Exam_20220817_Ali"
author: "Mohamed Ali - Mohal954"
date: "2023-05-23"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(123)
library(LaplacesDemon)
library(mvtnorm)
library(ggplot2)
source("ExamData.R")
```

# Question 1

## D and E

A function that computes the log posterior distribution

```{r q1de,echo=T}
logpost<- function(theta,x,n,a,b){
  loglik<- (3*n*log(theta))-((n)*log(2))-sum(log(x^4))-(theta*sum(1/x))
  logPrior<- (a*log(b))-log(factorial(a-1))+((a-1)*log(theta)-(b*theta))
  return(-sum(loglik+logPrior))
}

x=c(.7,1.1,.9,1.5)
n=length(x)
a=4
b=2
theta<-seq(0.01,.9,.01)
post_dist<-c()
for (i in 1:length(theta)) {
      post_dist[i]<-logpost(theta[i],x,n,a,b)
}
 post_plot<- data.frame(x=post_dist)
 ggplot() +
   geom_histogram(data=post_plot,aes(x = x,y=..density..),linetype=1,
                  fill='#14213D')+
   geom_density(data=post_plot,aes(x=x), color='#FCA311', size=1,
                fill='#FCA311',alpha=.5)+
   labs(title = "Posterior predictive distribution of theta ",
        x = "Theta",
        y = "Density") + theme_classic()
 
 

 initVal<-rep(0,length(theta))
OptimRes <- optim(initVal,logpost,gr=NULL,x,n,a,b,method=c("L-BFGS-B")
                   ,control=list(fnscale=-1),hessian=TRUE,lower = 0.1)
postmode<-as.matrix(OptimRes$par)

 
```

# Question 2

## A

Compute a 95% equal tail credible interval for beta_1 and interpret it

```{r q2a, echo=T}
# Drawing 20000 from the joint posterior we have 
mu_0=rep(0,3)
Sigma_0= 1/1e2*diag(3)
nIter= 20000
# Sample draw using the pre defined function
betaSample<-BayesLogitReg(y,X,mu_0,sigma_0,nIter)$betaSample

#a
# Finding 95% CI for beta_1
mean(betaSample[,2])
quantile(betaSample[,2],c(.05,.95))
```

Interpretion, as we can see from the results above the value of mean beta_1 one is lay between the  Lower and Upper pounds of the CI which shows a Sig level of the beta_1.

## B

Compute the joint posterior probability

```{r q2b, echo=T}
#b
# The Joint posterior probability for both beta_1 > 0 and beta_2 > 0

mean(betaSample[,2]>0)
mean(betaSample[,3]>0)

```


## C
Plot the posterior distribution

```{r q2c, echo=T}
#c
# We define the p_i as exp(b0+b1x1+b2x2)/1+ exp(b0+b1x1+b2x2
# Frist we define our new x
plot_fun<-function(intrcpt,x1,x2){
      X_new<-c(intrcpt,x1,x2)
      pr<-exp(betaSample%*%X_new)
      pi<- pr/(1+(pr))
      plot_data<- data.frame(x=pi,y=((1-pi)/pi))
      ggplot() +
        geom_histogram(data=plot_data,aes(x = x,y=..density..),linetype=1,
                       fill='#14213D')+
        geom_density(data=plot_data,aes(x=x), color='#FCA311', size=1,
                     fill='#FCA311',alpha=.5)+
        labs(title = "Posterior predictive distribution ",
             x = "Probs",
             y = "Density") + theme_classic()
}

## Example with the given data 
plot_fun(1,5,1)

## another example with different data
plot_fun(1,49,1)

```

As we can see from the both plots the calculating the probability for this bridge could be a litil bit tricky as it's not a full reliable, we can see that the values of the number of years since the bridge been construct have a mean value of 49 which make the new input value a bit unrealistic with our model data. thus the results we get is not resalable and we might need to question the rilability of this bridge data.  

## D

Plot the lower and upper limits of the posterior probability intervals as a function of x1

```{r q2d, echo=T}
simx_new<-as.matrix(data.frame(const=1,x1=seq(min(X[,2]),max(X[,2]),.1),x2=0))
pred<-betaSample%*%t(simx_new)
pi<-exp(pred)/1+exp(pred)
lower<-c()
upper<-c()
for(i in 1:ncol(pi)){
  upper[i]<-quantile(pi[,i],.95)
  lower[i]<-quantile(pi[,i],.05)
}

plt_df<- data.frame(x=seq(min(X[,2]),max(X[,2]),.1),CI_lower=lower,CI_upper=upper)

plt <- ggplot(plt_df,aes(x=x)) +
   geom_ribbon(aes(ymin = CI_lower, ymax = CI_upper)
              , alpha = 0.8,fill = "#EDC948")+
  labs(x = 'X grid', y = ' '
  ,title =' the lower and upper limits of the posterior probability intervals'
  ,color = "Line Legend")+
  theme(legend.position="bottom")
plt
```

## E

Consider a 40-year-old bridge i that was built using the certain technology. Plot the posterior distribution of pi and compute the posterior probability.

```{r q2e, echo=T}
plot_fun(1,40,1)
paste0("The posterior distribution of pi and compute the posterior probability that pi > 0.5 is  ,",round(mean(pi>.5),3))
```

# Question 3


```{r q3c, echo=T}
#Solution based on simulation

theta_new<-rbeta(1,38,22)
uti_buy<- 60*theta_new-20*(1-theta_new)
paste0("Utility Fucntion for day 51 buy ",round(uti_buy,3))
uti_notbuy<- 180*theta_new-240*(1-theta_new)
paste0("Utility Fucntion for day 51 not buy ",round(uti_notbuy,3))
```

Thus the company should not buy.