---
title: "Exam_20221020_Ali"
author: "Mohamed Ali - Mohal954"
date: "2023-05-23"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
source("ExamData.R")
```

# Question 1

## D

computes the log posterior distribution

```{r q1d,echo=T}
logpost<- function(x,N,n,a,b,theta){
  loglike<-sum(x)*log(theta) + (n*N-sum(x))*log(1-theta)
  logprior<- (a-1)*log(theta) + (b-1)*log(1-theta)
  return(-(loglike+logprior))
}

x<-c(13,8,11,7)
N=20
a=2
b=3
n=1
theta=seq(0.01,.99,.001)
theta_val<-c()
for (i in 1:length(theta)){
  theta_val[i]<-logpost(x,N,n,a,b,theta[i])
}


plot_df<- data.frame(x=theta_val)

ggplot(plot_df,aes(x = x)) +
  geom_histogram(aes(y=..density..),linetype=1,fill='#14213D')+
  geom_density(alpha=.2,color="#FCA311",size=1,fill="#FCA311")+
  labs(x = 'Probabilities', y = 'density',
       title ='Plot of the predictive distribution')
```

## E

You can also embed plots, for example:

```{r 1e, echo=T}
initVal=0.5
OptimRes <- optim(initVal,
                  logpost,
                  gr=NULL,
                  N=N,
                  x=x,
                  a=a,
                  b=b,
                  n=n,
                  method=c("L-BFGS-B"),
                  control=list(fnscale=-1),
                  lower = .2,
                  hessian=TRUE)

theta_val2<-dnorm(theta,mean=OptimRes$par,sd=sqrt(1/OptimRes$hessian))

plot(theta_val2)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

# Question 2

## A 

Plot the posterior distribution

```{r 2a, echo=T}
a=sum(Wolves$y)+40
b=length(Wolves$y)+2
plt_df<-data.frame(x=rgamma(10000,a,b))
paste0("the posterior probability is smaller than 21 = ",mean((rgamma(10000,a,b))<21))
ggplot(plt_df,aes(x = x)) +
  geom_histogram(aes(y=..density..),linetype=1,fill='#14213D')+
  geom_density(alpha=.2,color="#FCA311",size=1,fill="#FCA311")+
  labs(x = 'Probabilities', y = 'density',
       title ='Plot of the predictive distribution')
```

## B

```{r 2b, echo=T,warning=FALSE}
###
reg_a<- subset(Wolves,Wolves==1)
reg_b<- subset(Wolves,Wolves==0)

##### Infrance region A

aa=sum(reg_a$y)+40
ba=length(reg_a$y)+2
plt_df<-data.frame(x=rgamma(10000,aa,ba))

ggplot(plt_df,aes(x = x)) +
  geom_histogram(aes(y=..density..),linetype=1,fill='#14213D')+
  geom_density(alpha=.2,color="#FCA311",size=1,fill="#FCA311")+
  labs(x = 'Probabilities', y = 'density',
       title ='Plot of the predictive distribution')

##### Infrance region B

ab=sum(reg_b$y)+40
bb=length(reg_b$y)+2
plt_df<-data.frame(x=rgamma(10000,ab,bb))

ggplot(plt_df,aes(x = x)) +
  geom_histogram(aes(y=..density..),linetype=1,fill='#14213D')+
  geom_density(alpha=.2,color="#FCA311",size=1,fill="#FCA311")+
  labs(x = 'Probabilities', y = 'density',
       title ='Plot of the predictive distribution')

###### future week 
#a
theta_a<-rgamma(1,aa,ba)
regp_a<-c()
#b
theta_b<-rgamma(1,ab,bb)
regp_b<-c()

seq_<-seq(1,1000,1)
for (i in 1:length(seq_)) {
  regp_a[i]<-rpois(seq_[i],theta_a)
  regp_b[i]<-rpois(seq_[i],theta_b)
}
paste0("Probability that Yb>Ya is ",mean(regp_b>regp_a))
```


## C

we can prove this by calculating the mean number of wolvs in both regons and then we can compare the delt mean between the 2 regions. 
We can see that the avg number of wolves per week in region b is higher by 18% than region A.
```{r 2c, echo=T}
mean(regp_b)/mean(regp_a)-1
```


# Question 3

## A

```{r 3a, echo=T}
y=y
X=X
mu_0=rep(0,14)
Omega_0=1e2*diag(14)
v_0=1
sigma2_0=4^2
nIter=10000

## model
BayesLinReg(y, X, mu_0, Omega_0, v_0, sigma2_0, nIter)
beta_val<- BayesLinReg(y, X, mu_0, Omega_0, v_0, sigma2_0, nIter)$betaSample
## A


for (i  in 1:ncol(X)) {
  print(paste0("The 95% lower CI for beta_",i-1," is ", round(quantile(beta_val[,i],c(.05)),3), " and Upper CI is ",round(quantile(beta_val[,i],c(.95)),3)))
}

mean(beta_val[,2])
paste0("The 95% lower CI for beta_",i-1," is ", round(quantile(beta_val[,2],c(.05)),3), " and Upper CI is ",round(quantile(beta_val[,2],c(.95)),3))

```
the mean beat value lay between the upper and the lower CI which indecates that the beta is significat and the crim var have impact on the model.


## B

```{r 3b, echo=T}
sigma2_val<- BayesLinReg(y, X, mu_0, Omega_0, v_0, sigma2_0, nIter)$sigma2Sample

#mean
mean(sqrt(sigma2_val))

#median
median(sqrt(sigma2_val))
```

## C

```{r 3c, echo=T}

X_new<-as.matrix(data.frame(intercept=1,crim=seq(min(X[,2]),max(X[,2]),by=.1),zn=40,indus=1.5,
                     chas=0,nox=.5,rm=6,age=30,dis=5,rad=3,tax=300,ptratio=17,black=390,
                     lstat=4))

lower<-c()
upper<-c()
for (i in 1:nrow(X_new)) {
  res<- beta_val%*%X_new[i,]
  lower[i]<-quantile(res,c(.05))
  upper[i]<-quantile(res,c(.95))
}

plt_data<- data.frame(x=seq(min(X[,2]),max(X[,2]),by=.1),lower=lower,upper=upper)

plt <- ggplot(plt_data,aes(x=x)) +
  geom_ribbon(aes(ymin = lower, ymax = upper)
              , alpha = 0.8,fill = "#EDC948")+
  labs(x = 'x_1',title ='. Compute 95% equal tail posterior probability intervals'
       ,color = "Line Legend") +
  scale_color_manual(values = c("#14213D","#59A14F")
                     , labels = c("2","3"))+
  theme(legend.position="bottom")
plt
```

## D

```{r 3d, echo=T}
beta_val<- BayesLinReg(y, X, mu_0, Omega_0, v_0, sigma2_0, nIter)$betaSample

y_pred<-c()

for (i in 1:10000) {
  y_pred_val<-(beta_val[i,]%*%XNewHouse) + rnorm(1,mean = 0,sd=sqrt(sigma2_val[i]))
  y_pred[i]<-y_pred_val
}

hist(y_pred)
mean(y_pred>=20)
```
